A General and Parallel Platform for Mining Co-Movement Patterns over Large-scale Trajectories

### 2016-10-29 Camera-Ready/sec1_introduction.tex
Camera-Ready
Wait to be submitted

\section{Introduction}
The prevalence of positioning devices has drastically boosted 
the scale and spectrum of trajectory collection to an unprecedented level. 
Tremendous amounts of trajectories, in the form of sequenced spatial-temporal 
records, are continually generated from animal telemetry chips, 
vehicle GPSs and wearable devices. Data analysis on large-scale 
trajectories benefits a wide range of applications and services, 
including traffic planning~\cite{zheng2011urban}, animal analysis~\cite{li2010miningperiodic}, location-aware advertising~\cite{guo2016influence},  and social recommendations~\cite{bao2013survey}, to name just a few.

A crucial task of data analysis on top of trajectories is 
to discover co-moving objects. A \emph{co-movement} pattern~\cite{li2013managing,zheng2015survey} 
refers to a group of objects traveling together for a certain period of time 
and the group is normally determined by their spatial proximity. 
A pattern is prominent if the size of the group exceeds $M$ and the length of the duration exceeds $K$, where $M$ and $K$ are parameters specified by users. Rooted from such a basic definition 
and driven by different mining applications, there are many variants 
of co-movement patterns that have been developed with additional constraints.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
            \centering
            \includegraphics[width=\textwidth]{rw_perf_O.eps}
		\subcaption{ Varying No. of objects} 
    \label{fig:fig1}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
            \centering
            \includegraphics[width=\textwidth]{rw_perf_T.eps}
         \subcaption{ Varying No. of timestamps}
    \label{fig:fig2}
    \end{subfigure}
    %\vspace{-0.5em}
    \caption{Performance measures on existing co-movement patterns. A sampled GeoLife dataset
    is used with upto 2.4 million data points. Default parameters are $M=15$, $K=180$, $L=30$.}
    %\vspace{-0.5em}
    \label{fig:related_work_scalability}
\end{figure}

Therefore, in this paper, our contributions are to close these two gaps.
%In this paper, our contribution is towards closing these two gaps.
First, we propose the \emph{general co-movement pattern} (GCMP) which models
various co-moment patterns in a unified way and can avoid 
the \emph{loose-connection} anomaly. In GCMP,
we introduce a new gap parameter $G$ to pose a constraint on the temporal gap between two consecutive segments. 
%\revised{In GCMP, we adopt the concept of \emph{gap parameter $G$}
%in sequential pattern mining field~\cite{srikant1996mining} to constrain the temporal distance between two consecutive %segments.} 
By setting a feasible $G$, the loose-connection anomaly can be effectively controlled. In addition, our GCMP is also general. %and expressive. 
It can be reduced to any of the previous pattern by customizing its parameters.

Second, we investigate deploying our GCMP detector on the model MapReduce platform (i.e., Apache Spark) to tackle the scalability issue. Our technical contributions are three-fold. First, we design a baseline solution by replicating the snapshots 
%in multiple data chunks 
to support effective parallel mining. 
Second, we devise a novel \emph{Star Partitioning and ApRiori Enumerator} (SPARE) framework to resolve limitations of the baseline. 
SPARE achieves workload balance by partitioning objects into fine granular stars. 
For each partition, an Apriori Enumerator is adopted to mine the co-movement patterns. 
Third, we leverage the \emph{temporal monotonicity} property of GCMP 
to design several optimization techniques including \emph{sequence simplification}, \emph{monotonicity pruning} and \emph{forward closure check} to further reduce the number of candidates enumerated in SPARE.

>>>>

\section{Introduction}
The prevalence of positioning devices has drastically boosted 
the scale and spectrum of trajectory collection to an unprecedented level. 
Tremendous amounts of trajectories, in the form of sequenced spatial-temporal 
records, are continually being generated from animal telemetry chips, 
vehicle GPSs and wearable devices. Data analysis on large-scale 
trajectories benefits a wide range of applications and services, 
including traffic planning~\cite{zheng2011urban}, animal analysis~\cite{li2010miningperiodic}, location-aware advertising~\cite{guo2016influence},  and social recommendations~\cite{bao2013survey}, to name just a few.

A crucial task of data analysis on top of trajectories is 
to discover co-moving objects. A \emph{co-movement} pattern~\cite{li2013effective,zheng2015survey} 
refers to a group of objects traveling together for a certain period of time 
and the group is normally determined by their spatial proximity. 
A pattern is prominent if the size of the group exceeds $M$ and the length of the duration exceeds $K$, where $M$ and $K$ are parameters specified by users. Rooted from such a basic definition 
and driven by different mining applications, there are many variants 
of co-movement patterns that have been developed with additional constraints.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
            \centering
            \includegraphics[width=\textwidth]{rw_perf_O.eps}
		\subcaption{ Varying No. of objects} 
    \label{fig:fig1}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\textwidth}
            \centering
            \includegraphics[width=\textwidth]{rw_perf_T.eps}
         \subcaption{ Varying No. of timestamps}
    \label{fig:fig2}
    \end{subfigure}
    %\vspace{-0.5em}
    \caption{Performance measures on existing co-movement patterns. A sampled GeoLife dataset
    is used with upto 2.4 million data points. Default parameters are $M=15$, $K=180$, $L=30$.}
    %\vspace{-0.5em}
    \label{fig:related_work_scalability}
\end{figure}
%
%Therefore, in this paper, our contributions are closing
%these two gaps.
%
%to close these two gaps.
%
%Therefore, in this paper, our aim is to close these two gaps 
%via the following contributions.
%we aim to close these two gaps by making two contributions. 
In this paper, we resolve these two gaps by making the following contributions.
First, we propose the \emph{general co-movement pattern} (GCMP) which models
various co-moment patterns in a unified way and can avoid 
the \emph{loose-connection} anomaly. In GCMP,
we introduce a new gap parameter $G$ to pose a constraint on the temporal gap between two consecutive segments. 
By setting a feasible $G$, the loose-connection anomaly can be effectively controlled. In addition, our GCMP is also general. It can be reduced to any of the previous pattern by customizing its parameters.

Second, we investigate deploying our GCMP detector on the model MapReduce platform (i.e., Apache Spark) to tackle the scalability issue. Our technical contributions are threefold. First, we design a baseline solution by replicating the snapshots 
%in multiple data chunks 
to support effective parallel mining. 
Second, we devise a novel \emph{Star Partitioning and ApRiori Enumerator} (SPARE) framework to resolve limitations of the baseline. 
SPARE achieves workload balance by partitioning objects into fine granular stars. 
For each partition, an Apriori Enumerator is adopted to mine the co-movement patterns. 
Third, we leverage the \emph{temporal monotonicity} property of GCMP 
to design several optimization techniques including \emph{sequence simplification}, \emph{monotonicity pruning} and \emph{forward closure check} to further reduce the number of candidates enumerated in SPARE.

### 2016-10-29 Camera-Ready/sec5_spm.tex
Camera-Ready
Wait to be submitted

Our customized Apriori Enumerator is presented in Algorithm~\ref{algo:apriori_mining}. Initially, the edges (pairs of objects) in the star constitute the bottom level (Lines~\ref{code:init-start}-\ref{code:init-end}) and invalid candidates are excluded (Line~\ref{code:simp1}). An indicator \emph{level} is used to control the result size for candidate joins. During each iteration (Lines~\ref{code:level-start}-\ref{code:level-ends}), only candidates with object size equals to \emph{level} are generated (Line~\ref{code:join-start}). When two candidates $c_1$ and $c_2$ are joined, the new candidate becomes $c'=\langle c_1.O \cup c_2.O, c_1.T\cap c_2.T\rangle$ (Line~\ref{code:join}). To check the validity of the candidate, we calculate $\mathtt{sim}(c'.T)$. If its simplified sequence is empty, $c'$ is excluded from the next level (Line~\ref{code:simp2}). This ensures that all the candidates with $P.O\supseteq c'.O$ are pruned. If a candidate cannot generate any new candidate, then
it is directly reported (Lines~\ref{code:output1-start}-\ref{code:output1-end}).
To further improve the performance, we adopt the idea of \textit{forward closure}~\cite{wang2003closet+} and  aggressively check if the union of all the current candidates form a valid pattern (Lines~\ref{code:fc-checking-start}-\ref{code:fc-checking-end}). If yes, we can  terminate the algorithm early and output the results.

>>>>

Our customized Apriori Enumerator is presented in Algorithm~\ref{algo:apriori_mining}. Initially, the edges (pairs of objects) in the star constitute the bottom level (Lines~\ref{code:init-start}-\ref{code:init-end}) and invalid candidates are excluded (Line~\ref{code:simp1}). An indicator \emph{level} is used to control the result size for candidate joins. During each iteration (Lines~\ref{code:level-start}-\ref{code:level-ends}), only candidates with object size equals to \emph{level} are generated (Line~\ref{code:join-start}). When two candidates $c_1$ and $c_2$ are joined, the new candidate becomes $c'=\langle c_1.O \cup c_2.O, c_1.T\cap c_2.T\rangle$ (Line~\ref{code:join}). To check the validity of the candidate, we calculate $\mathtt{sim}(c'.T)$. If its simplified sequence is empty, $c'$ is excluded from the next level (Line~\ref{code:simp2}). This ensures that all the candidates with $P.O\supseteq c'.O$ are pruned. If a candidate cannot generate any new candidate, then
it is directly reported (Lines~\ref{code:output1-start}-\ref{code:output1-end}).
To further improve the performance, we adopt the idea of \textit{forward closure}~\cite{wang2003closet+,pei2000closet} and  aggressively check if the union of all the current candidates form a valid pattern (Lines~\ref{code:fc-checking-start}-\ref{code:fc-checking-end}). If yes, we can  terminate the algorithm early and output the results.
